---
title: 'Cleaning Final, Merged Dataset for Analysis'
output:
  html_document:
    df_print: paged
---

```{r include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

```{r}
library('data.table')
library('ggplot2')
library('stringr')
library('kableExtra')
```

```{r}
setwd('//cdc.gov/project/ATS_GIS_Store4/Projects/prj06135_Shigella_SVI/Data/FoodNet_NARMS')
dat<-read.csv('FN_NARMS_PulseNet_Census.csv', stringsAsFactors = F)
setDT(dat)
```

# Clean variables for demographic tables
****
## Age Range
AgeRange is derived from the **Age** variable. There are `r sum(is.na(dat$AgeRange))` records with missing age range.
However, there are additional age variables: **PatientAgeYears**, **PatientAgeMonths**, **PatientAgeDays**, and **PatientAgeFromPulsenet** so perhaps we can fill in some missing values. 
```{r, include=T}
t1<-table(rowSums(is.na(dat[, c('PatientAgeYears', 'PatientAgeMonths', 'PatientAgeDays', 'PatientAgeFromPulsenet')]))<4, is.na(dat$Age))
```
There are `r t1[4]` records that have a NULL Age value, but have at least one non-missing values for PatientAge* or PatientAgeFromPulesNet. 

Before we fill-in missing data, let's double-check to make sure that the original AgeRange was calculated correctly. 
```{r}
dat<-dat[order(PatientAgeYears),]
print(dat[AgeUnit=='Months', c('AgeRange', 'Age', 'AgeUnit', 'PatientAgeYears')] , nrow=25)
```
It looks like the age unit was neglected when calculating AgeRange. We will need to re-cut that variable.

It also appears that Age was calculated in years based on elapsed dates for a lot of records, those are missing an age unit. What I think happened is that AgeYrs, which was a FoodNet numeric double field calculated on elapsed time between DOB and DtSpec date, was combined with Age from NARMS which was a numeric integer that needs AgeUnit to be correctly interpreted. We will need to create an AgeClean field to fix this error. We will also use the PatientAgeFromPulseNet, PatientAgeYears, and PatientAgeMonths variables to fill in missing values.
```{r}
dat$AgeUnit[is.na(dat$AgeUnit) & !is.na(dat$Age)]<-'Years'
dat$PNAge<-ifelse(grepl("0-", dat$PatientAgeFromPulsenet)==T, 0, as.numeric(dat$PatientAgeFromPulsenet))
```

```{r}
dat$AgeClean<-floor(
               ifelse(dat$AgeUnit=='Years'   & !is.na(dat$Age), dat$Age,
                ifelse(dat$AgeUnit=='Months' & !is.na(dat$Age), dat$Age/12,
                 ifelse(is.na(dat$Age) & !is.na(dat$PNAge), dat$PNAge,
                   ifelse(is.na(dat$Age) & is.na(dat$PNAge) & !is.na(dat$PatientAgeYears), dat$PatientAgeYears, 
                     ifelse(is.na(dat$Age) & is.na(dat$PNAge) & is.na(dat$PatientAgeYears), dat$PatientAgeMonths/12, NA))))))
```

Ok, let's take a look at the new AgeClean variable and see how it compares to the original. We should see shift in cases original classified in the 5-24 ages to the <2 age ranges.  
```{r, echo=F}
hist(dat$Age, col=rgb(0,0,1,1/4), main='Original and New Age Variables', xlab='Age Variable')
hist(dat$AgeClean, col=rgb(1,0,0,1/4), add=TRUE)
legend("topright", c("Original", "New"), col=c(rgb(0,0,1,1/4), rgb(1,0,0,1/4)), lwd=10)

dat$AgeRangeNew<-cut(dat$AgeClean, breaks=c(0, 5, 18, 30, 40, 50, 199), right=F, 
                     labels=c('0-4','5-17','18-29','30-39','40-49','50+'))

dat<-dat[order(AgeClean),]
dat[!is.na(AgeRangeNew), list(MinAge=min(AgeClean, na.rm=T), MaxAge=max(AgeClean, na.rm=T)), by='AgeRangeNew']
```
There are now only `r sum(is.na(dat$AgeClean))` records with a missing AgeRangeNew value. Thankfully, the other demographic variables (gender, race, ethnicity) look good (no implausible or missing values).

## Group Serotypes
This code condenses the values in **FinalSpeciesName** into a smaller number of categories. This variable is included in the NARMS dataset. However, there are `r sum(is.na(dat$FinalSpeciesName))`records missing FinalSpeciesName values. There is another field (**serotypesummary**) that has species values that can be used to fill in missing values. The FoodNet Codebook states that the serotypesummary field is a CDC-cleaned version of SeroSite. Per the discussion with Naeemah Logan (5/27), the NARMS FinalSpeciesName field should take precedence. 

```{r}
dat$SpeciesClean<-ifelse(dat$FinalSpeciesName %in% c(NA, 'unknown'), 
                         tolower(word(dat$serotypesummary, 1)), dat$FinalSpeciesName)
dat$SpeciesClean[dat$SpeciesClean %in% c(NA, 'not')]<-'unknown' #adds not serotyped and not speciated to unknown
table(dat$FinalSpeciesName, dat$SpeciesClean, useNA='ifany')
```

## FoodNet Site
The State field is equivalent to FoodNet Site. 
```{r}
addmargins(table(dat$State, useNA ='ifany' ))
```

## Resistance Profile
For lab approved isolates not associated with an outbreak (Surveillance ==1 & Exclude ==0 & LabApproval ==1), we will categorize resistance. Note, the field **SurveillanceReportable** is a flag variable for cases with Surveillance ==1 & Exclude ==0 & LabApproval ==1, so I'm going to use that for brevity. 

Resistance is grouped into the following variables (not mutually exclusive):

Resistance Type | Definition     | New Field Name
--------------- | -------------- | --------------
Ampicillin-R    | AMP_Concl=='R' | AmpR
Azithromycin-R  | Azm_Concl=='R' | AzmR
Ciprofloxacin-R | CIP_Concl %in% c('I', 'R') | CipR
Trimethoprim-Sulfamethoxazole-R  | COT_Concl=='R' | CotR
Ceftriaxone-R   | AXO_Concl=='R' | AxoR
Multidrug-R     | AMP_Concl & AZM_Concl & CIP_Concl & COT_Concl ='R' | MDR
Extensive Drug-R | MDR==T & AXO_Concl=='R' | XDR

*Resistance does not include intermediate results, with the exception of the Ciprofloxacin-DSC category.* 
```{r}
#Individual categories of antibiotic resistance
dat[SurveillanceReportable ==1, c("AmpR", 'AzmR', "CipR", "CotR",  'AxoR') := 
      list(AMP_Concl=='R', 
           AZM_Concl=='R',
           CIP_Concl %in% c('I', 'R'),
           COT_Concl=='R', 
           AXO_Concl=='R')]

#Number of antibiotic resistance categories, multi-drug and x-drug resistant
abxlist<-paste0(c('AMP', 'AZM', 'CIP', 'COT', 'AXO'), "_Concl")
dat[SurveillanceReportable==1, AbxResSum :=rowSums(.SD=='R', na.rm=T), .SDcols=abxlist]

#Multi-drug resistance
dat[SurveillanceReportable==1, MDR := rowSums(.SD=='R', na.rm=T)==4, .SDcols=abxlist[1:4]]

#X-drug resistance
dat[SurveillanceReportable==1, XDR := MDR==T & AXO_Concl=='R']

#No drug resistance
dat[SurveillanceReportable==1, NoAbxRes := AbxResSum == 0]
```

There were a total of `r sum(dat$SurveillanceReportable, na.rm=T)` cases that had a primary purpose of surveillance, 
were lab approved and not excluded. These isolates are part of the NARMS surveillance scheme and are included in official NARMS reports.

Of those cases, there are a total of `r sum(dat$AbxResSum>2, na.rm=T)` cases that were resistant to three or more antibiotics, `r sum(dat$MDR==T, na.rm=T)` cases of multidrug resistance, and `r sum(dat$XDR==T, na.rm=T)` cases of extreme drug resistance. 

### Age density by antibiotic resistance category
```{r, echo=F}
temp<-rbind(data.frame(Age=na.exclude(dat$AgeClean[dat$AmpR==T]), Resistance='Ampicillin'),
            data.frame(Age=na.exclude(dat$AgeClean[dat$AzmR==T]), Resistance='Azithromycin'),
            data.frame(Age=na.exclude(dat$AgeClean[dat$CipR==T]), Resistance='Ciprofloxacin-R'),
            data.frame(Age=na.exclude(dat$AgeClean[dat$CotR==T]), Resistance='Cotrimoxazole'),
            data.frame(Age=na.exclude(dat$AgeClean[dat$AxoR==T]), Resistance='Ceftriaxone'),
            data.frame(Age=na.exclude(dat$AgeClean[dat$MDR==T]), Resistance='Multi-Drug'),
            data.frame(Age=na.exclude(dat$AgeClean[dat$XDR==T]), Resistance='XDR'),
            data.frame(Age=na.exclude(dat$AgeClean[dat$NoAbxRes==T]), Resistance='None'))

ggplot(temp, aes(x=Age, color=Resistance)) +
  geom_density(alpha=0.4, size=1.2)

setDT(temp)
t2<-temp[, list(N=.N, Min=min(Age), Mean=mean(Age), Median=median(Age), Max=max(Age)), by='Resistance']
t2 %>%
   kbl() %>%
   kable_minimal()
```

## Calculate measures of severity
### Fever
Fever include patients with a value of "YES" for Fever. There are `r sum(dat$Fever=="YES", na.rm=T)` cases with fever. 

### Hospitalization
First, we will verify that dates of admission or discharge to a hospital are within the surveillance period. In other words, we will make sure that these are all plausible dates.  
```{r}
dates<-c('DtAdmit', 'DtDisch', 'DtAdmit2', 'DtDisch2')
dat[, (dates) := lapply(.SD, as.Date, format="%m/%d/%Y"), .SDcols = c(dates)]
dat[, lapply(.SD, min, na.rm=T), .SDcols = c(dates)]
dat[, lapply(.SD, max, na.rm=T), .SDcols = c(dates)]
```

Looks good, so we will define hospitalization as any record with a value of YES for Hospital or HospTrans or a non-missing admit or discharge date (initial or second admit/discharge). 
```{r}
dat[, HospClean := Hospital=='YES'|HospTrans=='YES'|rowSums(is.na(.SD), na.rm=T)<4, .SDcols=c(dates)]
```
There are `r sum(dat$HospClean==T)` records that were hospitalized. 

### Bloody diarrhea
Bloody diarrhea includes records with a value of "YES" for BloodyDiarr. 
There are `r sum(dat$BloodyDiarr=="YES", na.rm=T)` cases with bloody diarrhea. 

### Bacteremia
There are three fields for specimen source (SpecimenSource, SourceSite, and SpecSrce) which have a "Blood" category.  The SpecSrce field is the CDC-derived field and is the most complete. 
```{r}
dat$Bacteremia<-ifelse(dat$SpecimenSource=='Blood'|
                       (is.na(dat$SpecimenSource) & dat$SpecSrce=='BLOOD') |
                       (is.na(dat$SpecimenSource) & is.na(dat$SpecSrce) & dat$SourceSite=='Blood'), T, F)
```
There are `r sum(dat$Bacteremia==T, na.rm=T)` cases with bacteremia. 

### Death
```{r}
dat$Death<-dat$Outcome=='DEAD'
```
There are `r sum(dat$Death==T, na.rm=T)` cases that died. 

### Non-Severe
**This was taken directly from the table shell footnote:**
*Criteria for non- severe classification are no fever, bloody diarrhea, and/or presence of hospitalization, bacteremia, or death.*

I'm not sure how to interpret the and/or specification, but I'm going to process this excluding those from the non-severe category since you can't get much more severe than death. This field is going to be mostly NA because Fever, BloodyDiarr and Bacteremia are mostly NA or UNKNOWN.
```{r}
dat$NonSevere<-ifelse(dat$Fever=='NO' & dat$HospClean==F & dat$BloodyDiarr=='NO' & dat$Bacteremia==F & dat$Death==F, T,
               ifelse(dat$Fever=='YES'| dat$HospClean==T| dat$BloodyDiarr=='YES' | dat$Bacteremia==T | dat$Death==T, F,
                      NA))
```
There are `r sum(dat$NonSevere==T, na.rm=T)` records that were classified as non-severe. 

# Select Fields for Analysis and Output
***
Due to the size of the merged dataset, we probably want to restrict columns to only those that will be used for analysis.
```{r}
options(scipen=500)
keep.fields<-c("ID",  "Year", 'SiteName',"CountyName" , "CTNO2000", "CTNO2010" ,
              "Race", "Ethnicity",  "Gender", 'AgeRangeNew', 'AgeClean', 
              "SpeciesClean", 
         Plymouth, United Kingdom     "AmpR", "CotR", "CipR", 'AxoR', 'AzmR', "AbxResSum", "MDR", "XDR",
              "Fever", "HospClean", "BloodyDiarr", "Bacteremia", "Death", "NonSevere")

#Output dataset: I'm not activating this code yet until I can confirm which fields we want to keep
#final<-dat[, ..keep.fields]
write.csv(dat, "//cdc.gov/project/ATS_GIS_Store4/Projects/prj06135_Shigella_SVI/Data/FoodNet_NARMS/analytic_file_V4.csv", row.names=F)
```

```{r}
temp<-read.csv("//cdc.gov/project/ATS_GIS_Store4/Projects/prj06135_Shigella_SVI/Data/FoodNet_NARMS/analytic_file_V4.csv", stringsAsFactors = F)
table(grepl('e', temp$CTNO2000))
```

