---
title: 'Cleaning Final, Merged Dataset for Analysis'

output:
  html_document:
    df_print: paged
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

```{r, include=F}
library('data.table')
library('ggplot2')
library('stringr')
library('kableExtra')
library('tidyverse')
library('readxl')
```

```{r}
#Import final FoodNet datatset
folder<-'//cdc.gov/project/ATS_GIS_Store4/Projects/prj06135_Shigella_SVI/Data/FoodNet_NARMS/'
dat<-read_excel(path=paste0(folder, 'loganshigella_20230413.xlsx'), col_types = "text")
setDT(dat)
```

# Data management
```{r}
#Missing values
dat[dat=='.']<-NA
dat$CTNO2000[dat$CTNO2000=='UNKNOWN']<-NA
dat$CTNO2010[dat$CTNO2010=='UNKNOWN']<-NA
```

```{r}
#numeric fields
numvars<-c('Year', 'Age')
dat[,  (numvars) := lapply(.SD, as.numeric), .SDcols=numvars]
```

# Census tract verification and imputation
```{r}
#Format census tract geoid's
dat[,'CTNO2000' :=  str_pad(CTNO2000, 11, 'left', pad=0)]
dat[,'CTNO2010' :=  str_pad(CTNO2010, 11, 'left', pad=0)]
```

This code is used to exclude non-valid census tract GEOID values and impute missing values based on both deterministic and stochastic methods. Missing 2000 census tract GEOID values are only imputed if the case was diagnosed before 2006 and missing 2010 values are only imputed if the case was diagnosed after 2005. Output includes summary statistics on the estimated accuracy of the imputation. 

```{r, include=FALSE}
source('census tract crosswalk and imputation.R')
```

## Data imputation results and quality

*Number of tracts with missing census tract information by year of diagnosis*

```{r}
missct
```

**Number of tracts and cases by type of census tract boundary change**
```{r}
tract_with_missing_by_rf
```

**Average percent of population contained within the imputed census tract by method.**
- Deterministic = Dtm
- Stoichastic = Stc. 
```{r, echo=F}
temp1
```

**Percent of tracts that had the same imputed tract result for both deterministic and stoichastic methods by year.**
```{r, echo=F}
temp2
```

**Estimated percent of cases assigned to correct tract by method and year (based on sample of tract with both values reported).**
```{r, echo=F}
temp4[order(Year),]
```

# Prepare and merge NARMS data

```{r}
source('narms_data_prep.R')
```

**Resistance is grouped into the following variables (not mutually exclusive):**

Resistance Type | Definition     | New Field Name
--------------- | -------------- | --------------
Ampicillin-R    | AMP_Concl=='R' | AmpR
Azithromycin-R  | Azm_Concl=='R' | AzmR
Ciprofloxacin-R | CIP_Concl %in% c('I', 'R') | CipR
Trimethoprim-Sulfamethoxazole-R  | COT_Concl=='R' | CotR
Ceftriaxone-R   | AXO_Concl=='R' | AxoR
Multidrug-R     | AMP_Concl & AZM_Concl & CIP_Concl & COT_Concl ='R' | MDR
Extensive Drug-R | MDR==T & AXO_Concl=='R' | XDR

*Resistance does not include intermediate results, with the exception of the Ciprofloxacin-R category.* 

```{r}
dat <- dat %>% 
  left_join(narms, join_by(SLabsID))
```

There are **`r sum(dat$SubPop==T, na.rm=T)`** cases that are had a primary purpose of surveillance, 
were lab approved, were not excluded, and have antibiotic resistance information. These isolates are 
part of the NARMS surveillance scheme and are included in official NARMS reports.

There are a total of `r sum(dat$AbxResSum>2, na.rm=T)` cases that were resistant to three or more antibiotics, `r sum(dat$MDR==T, na.rm=T)` cases of multidrug resistance, and `r sum(dat$XDR==T, na.rm=T)` cases of extensive drug resistance. 

# Demographic variables

## Age group

AgeGroup is derived from the **Age** field in the FoodNet dataset if available, and the **narms_age** field from the NARMS dataset if Age is missing.  

```{r}
agegroupfunc<-function(x)(as.numeric(cut(x, breaks=c(0, 5, 18, 30, 40, 50, 199), right=F)))
                                     
dat <- dat %>%
  mutate(AgeGroup=ifelse(is.na(Age), agegroupfunc(narms_age), agegroupfunc(Age)),
         AgeGroup=factor(AgeGroup, levels=1:6,labels=c('0-4','5-17','18-29','30-39','40-49','50+')))

dat$AgeGroup[is.na(dat$AgeGroup) & dat$AgeUnit=='Months']<-'0-4'
```

There are `r sum(is.na(dat$AgeGroup))` records with a missing AgeGroup value. 

## Sex
There are a few cases with unknown Sex that have a value for Gender and/or PatientSexFromPulseNet. 
```{r}
dat <- dat %>%
  mutate(Sex=ifelse(Sex==" ", "U", Sex), 
         Gender=ifelse(trimws(Gender)=="", "U", Gender),
         PatientSexFromPulsenet=ifelse(PatientSexFromPulsenet=="", "UNKNOWN", PatientSexFromPulsenet), 
         Sex=ifelse(Sex %in% c('U', NA) & Gender %in% c('M', 'F'), Gender, 
             ifelse(Sex %in% c('U', NA) & Gender == "U" & PatientAgeFromPulsenet %in% c('MALE', 'FEMALE'),
                    substr(PatientAgeFromPulsenet, 1, 1), Sex)))

```

There are `r sum(dat$Sex=='U')` case with an unknown sex. 

## Race and ethnicity

**Crosstab of race by ethnicity**
```{r}
table(dat$Race, dat$Ethnicity, useNA='ifany')
```

**Classfication of race and ethnicity with American Indian/Alaska Native and Pacific Islander and Native Hawaiian top-coded over ethnicity**
```{r}
dat <- dat %>%
  mutate(RacEthGroupA = factor(
           ifelse(Race=='W' & Ethnicity=='N', 1,
           ifelse(Race=='I', 2,
           ifelse(Race=='A' & Ethnicity=='N', 3, 
           ifelse(Race=='B' & Ethnicity=='N', 4, 
           ifelse(Race=='P', 5, 
           ifelse(Race=='M' & Ethnicity=='N', 6,
           ifelse(Race=='O' & Ethnicity=='N', 7,
           ifelse(Race %in% c('A', 'B', 'M', 'O', 'U', 'W') & Ethnicity=='H', 8, 9)))))))),
           levels = 1:9  ,
           labels=c('White-NH', 'AmInd-AKNat', 'Asian-NH', 'Black-NH', 'NatHwn-PI', 
                    'Multiracial-NH', 'Other-NH', 'Hispanic', 'Unknown')),
         RaceGroupB = factor(
           ifelse(Race=='W', 1, 
           ifelse(Race=='I', 2,
           ifelse(Race %in% c('A', 'P'), 3,
           ifelse(Race=='B', 4,
           ifelse(Race=='M', 5, 6))))), 
           levels = 1:6,
           labels = c('White', 'Am Ind-AK Native', 'Asian-Pac Isldr', 'Black',
                      'Mltracial', 'Otr and Unkwn')))
```

Number of cases within each race and ethnicity category
```{r}
dat %>% group_by(RacEthGroupA) %>% count()
```
Number of cases within each race category (does not classify by ethnicity)
```{r}
dat %>% group_by(RaceGroupB) %>% count()
```

# Disease severity

## Fever
```{r}
dat$Fever[is.na(dat$Fever)]<-'UNKNOWN'
```

Fever include patients with a value of "YES" for Fever. There are `r sum(dat$Fever=="YES", na.rm=T)` cases with fever. 

## Hospitalization
Verify that dates of admission or discharge to a hospital are within the surveillance period.
```{r}
dates<-c('DtAdmit', 'DtDisch', 'DtAdmit2', 'DtDisch2')
dat[, (dates) := lapply(.SD, function(x) as.Date(as.numeric(x), origin='1899-12-30')), .SDcols = c(dates)]
dat[, lapply(.SD, min, na.rm=T), .SDcols = c(dates)]
dat[, lapply(.SD, max, na.rm=T), .SDcols = c(dates)]
```

Define hospitalization as any record with a value of YES for Hospital or HospTrans or a non-missing admit or discharge date (initial or second admit/discharge). 
```{r}
dat[, HospClean := Hospital=='YES'|HospTrans=='YES'|rowSums(is.na(.SD), na.rm=T)<4, .SDcols=c(dates)]
```

There are `r sum(dat$HospClean==T)` cases that were hospitalized. 

## Bloody diarrhea
Bloody diarrhea includes records with a value of "YES" for BloodyDiarr. 
```{r}
dat$BloodyDiarr[is.na(dat$BloodyDiarr)]<-'UNKNOWN'
```

There are `r sum(dat$BloodyDiarr=="YES", na.rm=T)` cases with bloody diarrhea. 

## Bacteremia
There are three fields for specimen source (SpecimenSource, SourceSiteFromPulsenet, and SpecSrce) which have a "Blood" category.  The SpecSrce field is the CDC-derived field and is the most complete. 
```{r}
dat <- dat %>%
  mutate(Bacteremia= 
           ifelse(SpecimenSource=='Blood' | SpecSrce=='BLOOD'| SourceSiteFromPulsenet=='Blood', T, F))
```
There are `r sum(dat$Bacteremia==T, na.rm=T)` cases with bacteremia. 

## Death
```{r}
dat$Death<-dat$Outcome=='DEAD'
```
There are `r sum(dat$Death==T, na.rm=T)` cases that died. 

## Overall Severity (1+ Severity Classifications)
```{r}
dat$Severe<-dat$Fever=='YES'| dat$HospClean==T | dat$BloodyDiarr=='YES'| dat$Bacteremia==T | dat$Death==T
```
There are `r sum(dat$Severe==T, na.rm=T)` cases were classified as severe. 

### Output final analytic dataset
```{r}
options(scipen=500)
write.csv(dat, paste0(folder, "analytic_file_final_5192023.csv"), row.names=F)
```

