---
title: "Shigella Cases by Serverity and Demographic Characteristics"
author: "Sarah Rockhill"
project:
  type: website
  output-dir: docs
format: 
  html: 
    echo: false
    page-layout: full
    toc: true
---

These data include Shigella cases within the FoodNet Surveillance area that were diagnosed from 2004 through 2019. Cases with an unknown census tract of residence were excluded. Criteria for severe classification are presence of one or more of the following: fever, bloody diarrhea, hospitalization, bacteremia, or death.

```{ojs}
//| echo: false
data = FileAttachment("dem_1_data.csv").csv()
filtered = data.filter(function(rates) {
  return dem_cat.includes(rates.Category);
})

percents = FileAttachment("dem_1_percents.csv").csv({typed: true})
PercentsFiltered = percents.filter(function(percents) {
  return dem_cat.includes(percents.Category);
})
```

```{r, warning=F, echo=F}
require(data.table)
trends <-setDT(read.csv('dem_2_trends.csv', stringsAsFactors = F))
trends<-melt(trends, 
               id.vars=c("Category", "Values", "Year"),
               measure.vars=c("Cases", "Percent", "Adjusted.Rate"),
               variable.name="Type",
               value.name="Statistic")

ojs_define(dat=trends)
```

```{ojs}
//| echo: false
trendsFiltered = transpose(dat).filter(function(trends) {
    return dem_cat.includes(trends.Category) &&
    type_cat.includes(trends.Type);
})
```

## Select Demographic Category

```{ojs}
//| echo: false
//| panel: input
viewof dem_cat = Inputs.select(
  [ "Total",
    "Sex",
    "AgeGroup",
    "Ethnicity",
    "Race",
    "Species",
    "State",
    "Resistant to",
    "Antibiotic Resistance"
  ], 
  {value: ["Total"], label: html`<b>Demographic category</b>`}
)
```

### Shigella Cases by Severity

::: {layout="[ [2,1] ]"}
#### Age-Adjusted Rates

```{ojs}
//| echo: false
Plot.plot({
  marginLeft: 100,
  height: 375,
  width: 1000,
  grid: true,
  
  style: {
    fontFamily: "sans-serif",
    fontSize: "12px"
  },
  
  x: {
    grid: true,
    label: "Age-adjusted rate per 100K"},
  
  y: {
    label: null
  },
  
  facet: {
    data: filtered,
    y: "Type",
    marginRight: 90
  },
  
  fy: {
  label: null
  },
  
  marks: [
    Plot.frame(),
    Plot.barX(filtered, {x: "Rate", y: "Values", fill: "Values"}),
    Plot.ruleX([0])
  ]
})
```

#### Percent of Cases

```{ojs}
//| echo: false
Plot.plot({
  height: 200,
  color: {
    legend: true
  },
  y : {tickFormat: "s"},
  marks: [
    Plot.barY(PercentsFiltered, {
      x: "Values",
      y: "Percent",
      fill: "Type"
    }),
    Plot.ruleY([0])
  ]
})
```
:::

### Shigella Incidence Trends

```{ojs}
//| echo: false
//| panel: input
viewof type_cat = Inputs.radio(
  ["Adjusted.Rate", "Cases", "Percent"], 
  {value: "Adjusted.Rate", label: html`<b>Select statistic</b>`}
)
```

```{ojs}
//| echo: false
Plot.plot({
  style: "overflow: visible;",
  marginRight: 40,
  height: 400,
  width: 1000,
  color: {
    legend: true
    },
    
   y: {grid: true,
       label: "Type"},

  marks: [
    Plot.ruleY([0]), // freezing
    Plot.line(trendsFiltered, {
        x: "Year", 
        y: "Statistic", 
        stroke: "Values",
        marker: "circle"
        }),
    Plot.text(trendsFiltered, Plot.selectLast({
        x: "Year",
        y: "Statistic",
        z: "Values",
        text: "Values",
        textAnchor: "Year",
        dx: 10
      }))
      
      ]
})
```

## Shigella Incidence by State
```{ojs}
//| echo: false
//| panel: input
viewof map_cat = Inputs.radio(
  ["Adjusted.Rate", "Cases", "Percent"], 
  {value: "Adjusted.Rate", label: html`<b>Select statistic</b>`}
)

viewof map_year = Inputs.range(
  [2004, 2019], 
  {value: 2004, step: 1, label: html`<b>Year</b>`}
)
```

```{r, echo=F}
#Format data for mapping
library('tidycensus')
names(fips_codes)[names(fips_codes)=='state_code']<-'id'
stFips<-unique(fips_codes[, c('state', 'id')])

stateData<-merge(trends[Category=='State',], stFips, by.x='Values', by.y='state', all.x=T)
setnames(stateData, 'Statistic', 'value')

ojs_define(mapdat=stateData)
```

```{ojs}
//| echo: false
mapFiltered = transpose(mapdat).filter(function(stateData) {
    return map_cat.includes(stateData.Type) &&
           map_year == stateData.Year;
})

maxdomain = d3.max(mapFiltered, d => d.value)

```

```{ojs}
d3 = require("d3@6")
import { drawChoroplethMap } from "@clhenrick/reusable-choropleth-map"
import {Legend} from "@d3/color-legend"
```

```{ojs}
Legend(d3.scaleQuantize([0, maxdomain], d3.schemeYlGnBu[5]), {
  title: "Quintile"
})
```


```{ojs}
colorScale = d3
  .scaleQuantize()
  .domain([0, maxdomain])
  .range(d3.schemeYlGnBu[5])
  
tooltipContents = datum => {
  const {value } = datum;
  return `${value}`;
}
  
testWithStatesData = drawChoroplethMap({
  data: mapFiltered,
  colorScale,
  tooltipContents,
  adminLevel: "states",
  zoomEnabled: false
})
```



